---
title: "Analysis for academic reporting"
author: "Wenjun He"
date: "`r format(Sys.Date(),'%d %B, %Y')`"

output:
  html_notebook:
    toc: yes
---
# Background

The report of this analysis refer to two articles:
 - Prognostic and predictive value of a microRNA signature in adults
with T-cell lymphoblastic lymphoma
 - A CpG Methylation Classifier to Predict Relapse in Adults
with T-Cell Lymphoblastic Lymphoma

<br>

# Step 1 - Data preprocessing
```{r training_internal_testing and external set,echo=FALSE, warning=FALSE, message=FALSE}
library("readxl")


# training and internal test dataset
training_internal_testing_path <- "/hewenjun/rstudio_work/zhenweipeng_version3/training-1-20201104.xlsx" 
data_training_internal_testing <- read_excel(training_internal_testing_path)

variables_to_remove <- c()
for (i in 1:40){
  var_remove <- paste("miR",i,"rank",sep = "-")
  variables_to_remove <- c(variables_to_remove,var_remove)
}

variables_to_keep <- c()
for (c in colnames(data_training_internal_testing)) {
  if (! c %in% variables_to_remove){
    variables_to_keep <- c(variables_to_keep,c)
  }
}
data_training_internal_testing <- data_training_internal_testing[,variables_to_keep]

# splitting into training set and interval validation set
library(caret)
set.seed(12)
sample_value <-createDataPartition(y=data_training_internal_testing$Recurrence,p=0.7,list=FALSE)
train_data <- data_training_internal_testing[sample_value, ]
test_data <- data_training_internal_testing[-sample_value, ]


# validation test dataset
validation_data_path <- "/hewenjun/rstudio_work/zhenweipeng_version3/validation-20201224.xlsx" 
external_data <- read_excel(validation_data_path)
```

# Step 2 - Lasso cox model to build a risk score using miR variables
```{r lasso cox model,echo=FALSE, warning=FALSE, message=FALSE}
mir_variable <- c()
for (i in 1:40){
  var <- paste("miR",i,sep = "-")
  mir_variable <- c(mir_variable,var)
}

set.seed(8) # 12 1 can get 4 variables | 12 4 get 5 | 12 8 get 4 | 1 9 get 7
cv.fit <- cv.glmnet(data.matrix(train_data[,mir_variable]), 
                        Surv(as.numeric(as.matrix(train_data[,"RFS"])), as.numeric(as.matrix(train_data[,"Recurrence"]))), 
                        family="cox", 
                        type.measure="C" )

lasso_cox_fit <- glmnet(data.matrix(train_data[,mir_variable]), 
                            Surv(as.numeric(as.matrix(train_data[,"RFS"])), as.numeric(as.matrix(train_data[,"Recurrence"]))),
                            family="cox")

# Coefficients in lasso cox
Coefficients <- coef(lasso_cox_fit, s = cv.fit$lambda.min)

# Variables selected by lasso cox
selected_variables <- c()
Active.Index <- which(Coefficients != 0)
Active.Coefficients <- Coefficients[Active.Index]
for(j in mir_variable[Active.Index]){
  selected_variables <- c(selected_variables, j)
}
selected_variables

# coefficient plot of lasso cox
plot(cv.fit$glmnet.fit,xvar = "lambda",label = T)
abline(v=log(c(cv.fit$lambda.min,cv.fit$lambda.1se)),lty=2)

# tiff(file = "2.1-lasso.tiff", width = 1000, height = 500)
# plot(cv.fit$glmnet.fit,xvar = "lambda",label = T)
# abline(v=log(c(cv.fit$lambda.min)),lty=2)
# while (!is.null(dev.list()))  dev.off()

# Cv plot of lasso cox
plot(cv.fit)

# tiff(file = "2.2-lasso.tiff", width = 1000, height = 500)
# plot(cv.fit)
# while (!is.null(dev.list()))  dev.off()

# Risk score calculation: 
# Risk_score = miR-1*0.0149204144+miR-6*-0.0921783831+miR-9*-0.0461641026+miR-40*0.0003417941
train_data$risk_score <- train_data$`miR-1`*0.0149204144 + train_data$`miR-6`*-0.0921783831 + train_data$`miR-9`*-0.0461641026 + train_data$`miR-40`*0.0003417941

test_data$risk_score <- test_data$`miR-1`*0.0149204144 + test_data$`miR-6`*-0.0921783831 + test_data$`miR-9`*-0.0461641026 + test_data$`miR-40`*0.0003417941

external_data$risk_score <- external_data$`miR-1`*0.0149204144 + external_data$`miR-6`*-0.0921783831 + external_data$`miR-9`*-0.0461641026 + external_data$`miR-40`*0.0003417941
```
# Step 3 - Clinicopathological parameters of risk score
```{r prepare risk score variable ,echo=FALSE, warning=FALSE, message=FALSE}
# Using cut off the get the high risk and low risk
# Traning data
train.cut.RFS <- surv_cutpoint(train_data, time = "RFS", event = "Recurrence",
   variables = c("risk_score"))
summary(train.cut.RFS)
plot(train.cut.RFS, "risk_score", palette = "npg")
train.cut.RFS <- surv_categorize(train.cut.RFS)
train_data$risk_score_cat <- train.cut.RFS$risk_score

# Internal test data
test.cut.RFS <- surv_cutpoint(test_data, time = "RFS", event = "Recurrence",
   variables = c("risk_score"))
summary(test.cut.RFS)
plot(test.cut.RFS, "risk_score", palette = "npg")
test.cut.RFS <- surv_categorize(test.cut.RFS)
test_data$risk_score_cat <- test.cut.RFS$risk_score

# External test data
external.cut.RFS <- surv_cutpoint(external_data, time = "RFS", event = "Recurrence",
   variables = c("risk_score"))
summary(external.cut.RFS)
plot(external.cut.RFS, "risk_score", palette = "npg")
external.cut.RFS <- surv_categorize(external.cut.RFS)
external_data$risk_score_cat <- external.cut.RFS$risk_score
```

```{r clinicopathological,echo=FALSE, warning=FALSE, message=FALSE}
library(summarytools)
# Training data
st_options(plain.ascii = FALSE)
var_factor <- c("Death","Recurrence","Chemotherapy","Gender","Age","Symptom","BMI","Size","Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion","risk_score_cat")
for (var in var_factor){
  train_data[,var] <- as.factor(as.matrix(train_data[,var]))
}
dfSummary(train_data[,var_factor], valid.col = FALSE, graph.magnif = 0.75)
train_data[,var_factor] %>% group_by(risk_score_cat) %>% dfSummary()

# Test data
for (var in var_factor){
  test_data[,var] <- as.factor(as.matrix(test_data[,var]))
}
dfSummary(test_data[,var_factor], valid.col = FALSE, graph.magnif = 0.75)
test_data[,var_factor] %>% group_by(risk_score_cat) %>% dfSummary()

# External data
for (var in var_factor){
  external_data[,var] <- as.factor(as.matrix(external_data[,var]))
}
dfSummary(external_data[,var_factor], valid.col = FALSE, graph.magnif = 0.75)
external_data[,var_factor] %>% group_by(risk_score_cat) %>% dfSummary()
```
```{r chi square test,echo=FALSE, warning=FALSE, message=FALSE}
chiq_circulation <- function(var, dataset){
  print(var)
  mytable<-table(unlist(dataset[,var]),dataset$risk_score_cat)
  print(addmargins(mytable,2)) 
  print(prop.table(mytable,1),digits = 3)
  print(chisq.test(mytable))
  fisher.test(mytable)
}

# Training data
print("----------------------------Training data----------------------------")
for(var in var_factor){
  chiq_circulation(var, train_data)
}

# Test data
print("----------------------------Test data----------------------------")
for(var in var_factor){
  chiq_circulation(var, test_data)
}

# External data
print("----------------------------External data----------------------------")
for(var in var_factor){
  chiq_circulation(var, external_data)
}

```

# Step 4 - Validation for the risk score including: histogram, ROC, KM curves
```{r histogram,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)

histogram_risk <- function(data_set){
  data_set <- data_set[order(data_set$risk_score),]
  p = ggplot()+
  geom_histogram(aes(x=1:length(data_set$risk_score),y=data_set$risk_score,fill=data_set$Recurrence),stat='identity')+
  theme_minimal()+
  theme_classic()+
  theme(legend.position=c(0.2,0.8),plot.title = element_text(hjust = 0.5), legend.title = element_blank())+
  labs(title="Risk score for every patient", x="Weight", y="Risk score for four-miRNA-based classifier") +
  scale_fill_discrete(labels = c("Recurrence", "Not Recurrence"))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
         axis.line.x=element_blank())
  print(p)
  # tiff(file = "histgram-data.tiff", width = 1000, height = 500)
  # print(p)
  # while (!is.null(dev.list()))  dev.off()
}

# Training data
histogram_risk(train_data)
# Test data
test_data$Recurrence <- as.factor(test_data$Recurrence)
histogram_risk(test_data)
# External data
external_data$Recurrence <- as.factor(external_data$Recurrence)
histogram_risk(external_data)
```

```{r KM,echo=FALSE, warning=FALSE, message=FALSE}
generate_KM <- function(data_set){
  data_set$Recurrence <- as.integer(data_set$Recurrence)
res_cox<-coxph(Surv(RFS, Recurrence) ~risk_score_cat, data = data_set)
p <- ggsurvplot(survfit(Surv(RFS, Recurrence) ~risk_score_cat, data = data_set), data = data_set, 
           risk.table = TRUE,
           conf.int = TRUE,
           pval = TRUE,
           surv.median.line = "hv",
           pval.coord=c(45,0.65),
           pval.size=4)
p$plot = p$plot + 
ggplot2::annotate("text",
                    x = 50, 
                    y = 0.85,
                    label = paste("HR :",round(summary(res_cox)$conf.int[1],2))) + ggplot2::annotate("text",
                  x = 50, 
                  y = 0.75,
                  label =paste("(","95%CI:",round(summary(res_cox)$conf.int[3],2),"-",round(summary(res_cox)$conf.int[4],2),")",sep = "")) + labs(title="RFS survival analysis")
print(p)

# tiff(file = "KM.tiff", width = 1000, height = 500)
# print(p)
# while (!is.null(dev.list()))  dev.off()

}

# Training data
generate_KM(data_set = train_data)

# Testing data
generate_KM(data_set = test_data)

# External data
generate_KM(data_set = external_data)
```

```{r ROC,echo=FALSE, warning=FALSE, message=FALSE}
library(timeROC)

dd <- datadist(train_data)
options(datadist="dd")

generate_roc <- function(train_data){
  dd <- datadist(train_data)
  options(datadist="dd")
  roc_lasso <- timeROC(T=train_data$RFS, delta = train_data$Recurrence, marker=train_data$risk_score, cause=2, weighting = "marginal", times = c(3,6,12), ROC = TRUE, iid = TRUE)
  # confint(roc_lasso)
  plot(roc_lasso, time=3, col = "red")
  plot(roc_lasso, time=6, add=TRUE, col = "blue")
  plot(roc_lasso, time=12,add=TRUE, col = "brown")
  #plot(roc_lasso, time=24, add=TRUE, col = "cyan")
  legend("bottomright",c("ROC at 3 months","ROC at 6 months","ROC at 12 months"),col=c("red","blue","brown"),lty=1,lwd=1,text.width = 0.2, cex=0.7)
  
  # tiff(file = "ROC.tiff", width = 500, height = 500)
  # plot(roc_lasso, time=3, col = "red")
  # plot(roc_lasso, time=6, add=TRUE, col = "blue")
  # plot(roc_lasso, time=12,add=TRUE, col = "brown")
  # legend("bottomright",c("ROC at 3 months","ROC at 6 months","ROC at 12 months"),col=c("red","blue","brown"),lty=1,lwd=1,text.width = 0.2, cex=0.7)
  # while (!is.null(dev.list()))  dev.off()
}

generate_roc(train_data = train_data)
generate_roc(train_data = test_data)
generate_roc(train_data = external_data)
```
# Step 5 - model selection for RFS and OS
```{r cox model evaluation,echo=FALSE, warning=FALSE, message=FALSE}
# Merge three dataset
my.variable.list <- c("OS","Death","Recurrence","RFS", "Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion","risk_score")
#train_data$risk_score_cat <- as.character(train_data$risk_score_cat)
merge_data <- rbind(train_data[,my.variable.list], test_data[,my.variable.list])
merge_data <- rbind(merge_data, external_data[, my.variable.list])

# Get the cut off in total cases
merge.cut.RFS <- surv_cutpoint(merge_data, time = "RFS", event = "Recurrence",
   variables = c("risk_score"))
plot(merge.cut.RFS, "risk_score", palette = "npg")
merge.cut.RFS <- surv_categorize(merge.cut.RFS)
merge_data$risk_score_cat <- merge.cut.RFS$risk_score


merge_data$risk_score_cat <- as.integer(as.factor(merge_data$risk_score_cat))
merge_data$risk_score_cat[merge_data$risk_score_cat==2] <- 0
library(My.stepwise)
my.variable.list.cox_to_integer <- c("Death","Recurrence","Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion")
for(var in my.variable.list){
  merge_data[,var] <- as.integer(as.matrix(unlist(merge_data[,var])))
}

my.variable.list.cox <- c("Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion","risk_score_cat")
# RFS model
My.stepwise.coxph(Time = "RFS", Status = "Recurrence", variable.list = my.variable.list.cox, data = merge_data)
# OS model
My.stepwise.coxph(Time = "OS", Status = "Death", variable.list = my.variable.list.cox, data = merge_data)
```

# Step 6 - Prediction of the benefits from Chemotherapy
```{r KM curves,echo=FALSE, warning=FALSE, message=FALSE}
generate_KM_Chemotherapy_RFS <- function(data_set){
  data_set$Recurrence <- as.integer(data_set$Recurrence)
res_cox<-coxph(Surv(RFS, Recurrence) ~Chemotherapy, data = data_set)
p <- ggsurvplot(survfit(Surv(RFS, Recurrence) ~Chemotherapy, data = data_set), data = data_set, 
           risk.table = TRUE,
           conf.int = TRUE,
           pval = TRUE,
           surv.median.line = "hv",
           pval.coord=c(32,0.65),
           pval.size=4)
p$plot = p$plot + 
ggplot2::annotate("text",
                    x = 35, 
                    y = 0.75,
                    label = paste("HR :",round(summary(res_cox)$conf.int[1],2))) + ggplot2::annotate("text",
                  x = 35, 
                  y = 0.70,
                  label =paste("(","95%CI:",round(summary(res_cox)$conf.int[3],2),"-",round(summary(res_cox)$conf.int[4],2),")",sep = "")) + labs(title="RFS survival analysis")
print(p)

# tiff(file = "KM_RFS.tiff", width = 1000, height = 500)
# print(p)
# while (!is.null(dev.list()))  dev.off()

}

generate_KM_Chemotherapy_OS <- function(data_set){
  data_set$Death <- as.integer(data_set$Death)
res_cox<-coxph(Surv(OS, Death) ~Chemotherapy, data = data_set)
p <- ggsurvplot(survfit(Surv(OS, Death) ~Chemotherapy, data = data_set), data = data_set, 
           risk.table = TRUE,
           conf.int = TRUE,
           pval = TRUE,
           surv.median.line = "hv",
           pval.coord=c(32,0.65),
           pval.size=4)
p$plot = p$plot + 
ggplot2::annotate("text",
                    x = 35, 
                    y = 0.75,
                    label = paste("HR :",round(summary(res_cox)$conf.int[1],2))) + ggplot2::annotate("text",
                  x = 35, 
                  y = 0.70,
                  label =paste("(","95%CI:",round(summary(res_cox)$conf.int[3],2),"-",round(summary(res_cox)$conf.int[4],2),")",sep = "")) + labs(title="OS survival analysis")
print(p)
# tiff(file = "KM_OS.tiff", width = 1000, height = 500)
# print(p)
# while (!is.null(dev.list()))  dev.off()

}

# RFS
# All cases
generate_KM_Chemotherapy_RFS(data_set = merge_data)
# All cases + miRNA-defined-low-risk
generate_KM_Chemotherapy_RFS(data_set = merge_data[merge_data$risk_score_cat==0,])
# All cases + miRNA-defined-high-risk
generate_KM_Chemotherapy_RFS(data_set = merge_data[merge_data$risk_score_cat==1,])

# OS
# All cases
generate_KM_Chemotherapy_OS(data_set = merge_data)
# All cases + miRNA-defined-low-risk
generate_KM_Chemotherapy_OS(data_set = merge_data[merge_data$risk_score_cat==0,])
# All cases + miRNA-defined-high-risk
generate_KM_Chemotherapy_OS(data_set = merge_data[merge_data$risk_score_cat==1,])
```

```{r forest plot,echo=FALSE, warning=FALSE, message=FALSE}
library(forestplot)

cochrane_from_rmeta <- 
  structure(list(
    mean  = c(NA,NA,0.55, NA,0.56, 0.62,NA,0.68,NA,0.57,0.82), 
    lower = c(NA,NA,0.44,NA,0.32,0.48,NA,0.54,NA,0.33,0.63),
    upper = c(NA,NA,0.7,NA,0.98,0.8,NA,0.86,NA,0.99,1.07)),
    .Names = c("mean", "lower", "upper"), 
    row.names = c(NA, -11L), 
    class = "data.frame")

tabletext <- cbind(
  c("Subgroup", "RFS", "All cases", "Four-miRNA-based", 
    "  Low risk", "  High risk", "OS", "All cases", "Four-miRNA-based", 
    "  Low risk", "  High risk"),
  c("End point", " ", "0.55(0.44-0.7)", " ", 
    "0.56(0.32-0.98)", "0.62(0.48-0.8)", " ", "0.68(0.54-0.86)", " ", 
    "0.57(0.33-0.99)", "0.82(0.63-1.07)"),
  c("P value", " ", "<0.0001", " ", 
    "0.037", "0.00011", " ", "0.0011", " ", 
    "0.042", "0.14"))
forestplot(tabletext, 
           graph.pos = 2,
           cochrane_from_rmeta,new_page = TRUE,
           #is.summary = c(TRUE,TRUE,rep(FALSE,8),TRUE),
           clip = c(0.1,2.5), 
           xlog = TRUE, 
           col = fpColors(box = "royalblue",
                          line = "darkblue",
                          summary = "royalblue"))
# tiff(file = "forestplot.tiff", width = 500, height = 500)
# forestplot(tabletext, 
#            graph.pos = 2,
#            cochrane_from_rmeta,new_page = TRUE,
#            #is.summary = c(TRUE,TRUE,rep(FALSE,8),TRUE),
#            clip = c(0.1,2.5), 
#            xlog = TRUE, 
#            col = fpColors(box = "royalblue",
#                           line = "darkblue",
#                           summary = "royalblue"))
# while (!is.null(dev.list()))  dev.off()
```
# Step 7 - Nomogram plot
```{r cox in training set,echo=FALSE, warning=FALSE, message=FALSE}
library(survival)
library(rms)


my.variable.list.train <- c("OS","Death","Recurrence","RFS", "Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion","risk_score_cat")
training_cox_data <- train_data[,my.variable.list.train]

training_cox_data$risk_score_cat <- as.integer(as.factor(training_cox_data$risk_score_cat))
training_cox_data$risk_score_cat[training_cox_data$risk_score_cat==2] <- 0

my.variable.list.cox_to_integer <- c("Death","Recurrence","Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion")
for(var in my.variable.list.cox_to_integer){
  training_cox_data[,var] <- as.integer(as.matrix(unlist(training_cox_data[,var])))
}

my.variable.list.cox <- c("Chemotherapy", "Gender","Age", "Symptom", "BMI", "Size", "Lymphnode","TBIL","ALT","GGT","CA199","CEA","Vascularinvasion","risk_score_cat")
My.stepwise.coxph(Time = "RFS", Status = "Recurrence", variable.list = my.variable.list.cox, data = training_cox_data)

training_cox_data$risk_score_cat <- as.factor(training_cox_data$risk_score_cat)
training_cox_data$Chemotherapy <- as.factor(training_cox_data$Chemotherapy)
training_cox_data$CEA <- as.factor(training_cox_data$CEA)
training_cox_data$Age <- as.factor(training_cox_data$Age)

```

```{r Nomogram plot,echo=FALSE, warning=FALSE, message=FALSE}
dd<-datadist(training_cox_data)
options(datadist="dd")
f <- cph(Surv(RFS, Recurrence)~risk_score_cat+Chemotherapy+CEA+Age, data = training_cox_data, x=TRUE, y=TRUE, surv = TRUE)
survival <- Survival(f)
survival1 <- function(x)survival(3,x)
survival2 <- function(x)survival(6,x)
survival3 <- function(x)survival(12,x)
nom2 <- nomogram(f, fun = list(survival1, survival2,survival3),
                fun.at = c(0.01, seq(0.1,0.9,by=0.05), 0.99),
                funlabel = c('3 months survival','6 months survival', '1 year survival'))
plot(nom2)

# tiff(file = "nomogram.tiff", width = 800, height = 500)
# plot(nom2)
# while (!is.null(dev.list()))  dev.off()
```